---
title: "Conformal Prediction Intervals"
output: html_notebook
date: "2021-04-01"
---

# General useful resources

- https://github.com/valeman/awesome-conformal-prediction
- https://github.com/ryantibs/conformal



There is a [difference](https://robjhyndman.com/hyndsight/intervals/) between prediction and confidence intervals. 

# Prediction interval in general

- interval associated with a random variable that will be observed in future and the specific probability of the random variable being situated in this interval

Prediction interval for **new** response: 
$$
	\hat y_h \pm t_{(1-\alpha/2, n-2)} \cdot SE_{pred}
$$
	
$$
	SE_{pred} = \sqrt{MSE \cdot \left(1+\frac{1}{n}+\frac{(x_h - \bar x)^2}{\sum(x_i - \bar x)^2}\right)}
$$
		



Prediction interval always narrower due to "extra MSE" term. 

# Confidence interval in general

- frequentist concept concerned with the position of the non-random, but unknown parameter in relation to the interval ("containing the true parameter with some probability, given a large number of repeated samples")

Confidence interval for **mean** response: 

$$
		\hat y_h \pm t_{(1-\alpha/2, n-2)} \cdot SE_{fit}
$$


$$
		SE_{fit} = \sqrt{MSE \cdot \left(\frac{1}{n}+\frac{(x_h - \bar x)^2}{\sum(x_i - \bar x)^2}\right)}
$$

$SE_{fit}$ can approach zero 0 due to missing "extra MSE" term. 

# Bayesian credible interval

- interval associated with the posterior distribution of a parameter (parameters are random variables in Bayesian perspective). Similarity to prediction interval, but here the interval is concerned with a parameter value in contrast to the association with an observation when talking about prediction intervals. 



# Conformal prediction framework

* [Vovk et al. 2008](https://jmlr.org/papers/v9/shafer08a.html), more thorough treatment in ['Algorithmic  Learning in a Random World' by Vovk et al. (2005)](https://link.springer.com/book/10.1007/b106715). 



- Why? Validity of prediction intervals. The probability of excluding the correct target value is bounded by a predefined confidence level. It makes it possible to assess the uncertainty of each single prediction.


# Regression conformal prediction intervals

* [Johansson et al. 2014](https://link.springer.com/article/10.1007/s10994-014-5453-0
) 

## Inductive conformal prediction

3-Step-Procedure:

1. Divide training set into two disjoint sets
	- proper training set $Z^t$
	- calibration set $Z^C$
2. Train the model using proper training set
3. For each calibration instance: 
	- predict output
	- calculate non-conformity score $\alpha_i=|y_i - \hat y_i|$

4. Calculate the prediction region: 

$$
\hat Y_j^{\delta} = \hat y_j  \pm \alpha_{s(\delta)}
$$

with the smallest $\alpha_{s(\delta)} \in S$ satisfying the equation

$$
\frac{\#\{z_i \in Z^C | \alpha_i < \alpha_{s(\delta)}+1 \}}{|Z^C|+1} \geq 1-\delta
$$

Let's generate some data based on the Friedman-problem from the `mlbench`-package and explore the data in a principled way.  

```{r}

library(tidyverse)
library(reshape2)


dat_frame  <- mlbench::mlbench.friedman1(n = 1000, sd = 1) 
dat_x  <- dat_frame$x %>% 
	as_tibble()
dat_y <- dat_frame$y %>% 
	as_tibble()

# independent variables
summary(dat_x)

# dependent variable
summary(dat_y)

# boxplots
boxplot(dat_x)

# density plots of vars

#Convert wide to long
melt.x <- reshape2::melt(dat_x)

# plot densities
ggplot(data = melt.x, aes(x = value)) + 
	geom_histogram() + 
	facet_wrap(~variable, scales = "free")


```



Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

