---
title: "Treatment Effects"
author: "Lennard Maßmann"
date: "5/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lei et al. arxiv paper on conformal inference of ITEs and counterfactuals. 

# Individual Treatment Effect

$$
\tau_i = Y_i(T_i=1) - Y_i(T_i=0) 
$$

with

$$
Y^{obs} = 
\begin{cases}
  Y_i(1)  \text{, if } T_i = 1\\    
  Y_i(0)  \text{, if } T_i = 0
\end{cases}
$$




- ITE unobserved by definition as we only observe one of the two potential outcomes. 

```{r}

```


# Average Treatment Effect



# Conditional Average Treatment Effect

$$
\tau(x) = \mathbb{E}[Y_i(T_i=1) - Y_i(T_i=0)|X=x] 
$$

Under **strong ignorability**, one can show that: 

$$
\tau(x) = \mathbb{E}[Y^{obs}|X=x, T=1] - \mathbb{E}[Y^{obs}|X=x, T=0]
$$
Estimation problem reduces to that of estimating certain conditional expectations. Computing reliable confidence bands difficult (finite samples, high dimensions)

- Expectation of ITE conditional on covariate values

- more detailed summary than ATE but: variance conditional on covariates not interpreted (crucial if we assume that covariates do not explain away all variation in ITE)

- Concern: 
	+ variability of response around reg function
	+ variability of CATE estimators due to finite samples
	
Therefore, we need **at the same time**: 

- many covariates to justify enough variability in response
- perfect estimate of CATE for every value of covariate

Both unrealistic in real world setting. 


# Quantile Treatment Effect

Powell (2020) and Fort (2016). 

$$
\tau(q) = F^{-1}_{Y(1)}(q) - F^{-1}_{Y(0)}(q) \text{ , } q \in [0,1]
$$

See `qte`-package.

 




